Identifying Software Performance Changes Across Variants and Versions

### 2020-08-16 paper.tex
added symbols

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{mathtools} 
\usepackage{subcaption}
\usepackage{caption}
\usepackage{tikz}
\usepackage{mdframed}
\usepackage{wrapfig}
\usepackage{caption}
%\usepackage{tikz}
%\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
%		\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

@@

	Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points.
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A novel technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions} or archived at {\color{red}\url{https://archive.softwareheritage.org/browse/origin/https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions}}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}

@@

	To obtain accurate estimations with a limited budget of measurements, we require a different approach. We propose an iterative and adaptive sampling approach. That is, we use an initial, but small sample set to explore the problem space and then increase the level of granularity at promising regions and dimensions (development history segments as well as individual features and interactions, in our case). 
	Considering introductory DBMS example (cf. Figure~\ref{fig:dbms_performance}), we illustrate our sampling strategy in Figure~\ref{fig:sampling plan}. Here, each box depicts a possible measurements, i.e., pair of configuration and commit. The black bars represent change points hidden in the software system‘s performance histories, i.e., measurements immediately before and after a change point are dissimilar. The current state of the sample set comprises all measurements that are filled in grey.

	To include new measurements into the sample set, we take into account four possible objectives. On the one hand, our sample set might already hint to some possible change points and associated configuration options. On the other hand, our sample set might be too sparse and we cannot infer some change points yet. To address this trade-off, we split our objectives into exploitation and exploration.

	 For \emph{exploitation}, we sample new measurements to verify guesses based on the current state. That is, in Figure~\ref{fig:sampling plan}, we might include the measurements depicted by black-filled symbols. The black-filled stars represent measurements are of interest because we have already identified one change point for the configuration \textsc{Compression}. We exploit this knowledge to test for further configuration options involved in this change point. The black-filled pentagons represent measurements of interest, because we have identified another early change point for \textsc{Compression}, but within a rather broad range of commits. Thus, we include further measurements from that range to narrow down the possible range for this particular change point.

	 For \emph{exploration}, we might include the measurements by non-filled symbols. The stars represent measurements that explore new configurations. In our example, the interaction between \textsc{Compression} and \textsc{Encryption} is the only configuration left to be measured. We include several commits of a particular new configuration to unveil possible performance variation that indicate possible change points. The pentagons represent measurements that are of interest because they increase the overall measurement coverage. We include measurements from large intervals of not yet measured commits since possible change points can be hidden here.

	These four strategies (exploration and exploitation for configurations and commits) prioritize measurements to be included nex into the sample set. Adaptive sampling techniques have been successfully applied to obtain both performance-influence models~\cite{siegmundPredictingPerformanceAutomated2012,sarkarCostEfficientSamplingPerformance} and performance histories\cite{muhlbauer_accurate_2019} before. However, it is unclear whether fewer measurements are sufficient to assess the performance-influence of configuration options and interactions with respect to version changes. This is what we aim to accomplish with our work.

>>>>

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{mathtools} 
\usepackage{subcaption}
\usepackage{caption}
\usepackage{tikz}
\usepackage{mdframed}
\usepackage{wrapfig}
\usepackage{caption}

\usepackage{oPlotSymbl}
%\usepackage{tikz}
%\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
%		\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

@@

	Our main idea is as follows: We address the configuration complexity of this problem by selecting representative sample sets of configurations.
	Next, we uniformly sample a constant number of commits for each configuration and conduct respective performance measurements.
	Based on these measurement data, we assemble a prediction model.
	For each configuration, we estimate the likelihood of each commit being a change point, i.e. that performance of some configurations changes abruptly compared to the previous commit.
	Next, we leverage similarities in the performance histories of configurations that share common options.
	Since we sample the commits for each configuration independently, for each change point, we obtain many estimations using measurements of different commits.
	Overall, this allows us to obtain more accurate estimations of performance-changing commits with tractable effort.
	Based on a mapping from configurations to predicted change point probabilities for each configuration and commit, we derive the options responsible for each particular change point: configuration-dependent change points.
	In summary, we offer the following contributions:
	\begin{itemize}
		\item A novel technique to effectively identify shifts in performance of configurable software systems. We are able to pinpoint causative commits and affected configuration options with high accuracy and tractable effort.
		\item A feasibility demonstration of our approach by implementing an adaptive learning strategy to obtain accurate estimations with acceptable measurement cost.
		\item An evaluation using both synthetic and real-world performance data from three configurable software systems. Synthetic data lets us assess our model and approach conceptually and at scale, whereas we are able to assess practicality with real-world data.
		\item A companion website\footnote{\url{https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions} or archived at \url{https://archive.softwareheritage.org/browse/origin/https://github.com/AI-4-SE/Changepoints-Across-Variants-And-Versions}} providing supplementary material including a reference implementation of our approach, performance measurement data, and additional visualizations.
	\end{itemize}

@@

	To obtain accurate estimations with a limited budget of measurements, we require a different approach. We propose an iterative and adaptive sampling approach. That is, we use an initial, but small sample set to explore the problem space and then increase the level of granularity at promising regions and dimensions (development history segments as well as individual features and interactions, in our case).

	Considering the introductory DBMS example (cf. Figure~\ref{fig:dbms_performance}), we illustrate our sampling strategy in Figure~\ref{fig:sampling plan}. Here, each box depicts a possible measurement, i.e., a pair of a configuration and a commit. The black bars represent change points hidden in the software system‘s performance histories, i.e., measurements immediately before and after a change point are dissimilar. The current state of the sample set comprises all measurements that are filled in grey.

	To include new measurements into the sample set, we take into account four possible objectives. On the one hand, our sample set might already hint to some possible change points and associated configuration options. On the other hand, our sample set might be too sparse and we cannot infer some change points yet. To address this trade-off, we split our objectives into exploitation and exploration of configurations and commits, respectively.

	 For \emph{exploitation}, we sample new measurements to verify guesses based on the current state. That is, in Figure~\ref{fig:sampling plan}, we might include the measurements depicted by black-filled symbols~($\starletfill$/$\pentagofill$). The black-filled stars~($\starletfill$) represent measurements are of interest because we have already identified one change point for the configuration \textsc{Compression}. We exploit this knowledge to test for further configuration options involved in this change point. The black-filled pentagons~($\pentagofill$) represent measurements of interest, because we have identified another early change point for \textsc{Compression}, but within a rather broad range of commits. Thus, we include further measurements from that range to narrow down the possible range for this particular change point.

	 For \emph{exploration}, we might include the measurements by non-filled symbols~($\starlet$/$\pentago$). The non-filled stars~($\starlet$) represent measurements that explore new configurations. In our example, the interaction between \textsc{Compression} and \textsc{Encryption} is the only configuration left to be measured. We include several commits of a particular new configuration to unveil possible performance variation that indicate possible change points. The non-filled pentagons ($\pentago$) represent measurements that are of interest because they  increase the overall measurement coverage. We include measurements from large intervals of not yet measured commits since possible change points can be hidden here.

	These four strategies (exploration and exploitation for configurations and commits) prioritize measurements to be included next into the sample set. Adaptive sampling techniques have been successfully applied to obtain both performance-influence models~\cite{siegmundPredictingPerformanceAutomated2012,sarkarCostEfficientSamplingPerformance} and performance histories~\cite{muhlbauer_accurate_2019} before. However, it is unclear whether fewer measurements are sufficient to assess the performance-influence of configuration options and interactions with respect to version changes. This is what we aim to accomplish with our work.
